Accuracy and precision of zero-heat-flux temperature monitoring: a systematic review and meta- analysis \thanks{Grants or other notes about the article that should go on the front page should be placed here. General acknowledgments should be placed at the end of the article.}

### 2019-09-18 zhf_manuscript/zhf_manuscript.tex
prisma right size
dataframe ma output

@@

    \titlerunning{Short form of title (if too long for head)}

\author{  Aaron Conway \and  Âuthóř 2 \and  }

    \authorrunning{ Conway, }

\institute{
        Aaron Conway \at
     Peter Munk Cardiac Centre, UHN \& Lawrence S. Bloomberg Faculty of
 Nursing, University of Toronto \\
     \email{\href{mailto:aaron.conway@utoronto.ca}{\nolinkurl{aaron.conway@utoronto.ca}}}  %  \\
%             \emph{Present address:} of F. Author  %  if needed
    \and
        Âuthóř 2 \at
     Department of ZZZ, University of WWW \\
     \email{\href{mailto:djf@wef}{\nolinkurl{djf@wef}}}  %  \\
%             \emph{Present address:} of F. Author  %  if needed
    \and
    }

@@

Zero-heat-flux (ZHF) thermometry is an alternative non-invasive method
that allows for continuous monitoring of core temperature. Originally
developed in the 1970s, ZHF technology was recently implemented in the
3M SpotOn (3M, St Paul, MN), a single-use, disposable sensor \((ref)\).
In practice, the ZHF sensor, such as the 3M SpotOn, is placed on the
lateral surface of the forehead and initially warmed to equilibrate the
temperature of the skin surface to the underlaying core
tissues\((ref)\). Equipped with a thermal insulator, the ZHF sensor
eliminates heat loss to the environment to allow for changes in core
temperature to be directly reflected by a change in skin surface
temperature \((ref)\).

The agreement between ZHF and established core temperature devices has
been investigated in several studies conducted within the clinical
setting, each with varied results. Appraisal of these studies and
synthesis of the results would aid clinicians in deciding the
appropriate circumstances in which ZHF may be used.

We aimed to determine if ZHF has clinically acceptable accuracy and
precision relative to established core temperature measurement devices.
Accuracy is defined as the average difference between temperature
measurements from the ZHF and comparator device and precisions as the
variance (typically reported as the SD) in the differences.

@@

A systematic review was conducted in accordance with a predetermined
protocol (PROSPERO trial registration number:
\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#).

@@

Published studies were found by searching Medline and EMBASE from
January 2000 to July 2019, \textgreater{}as well as the reference lists
of articles identified to be relevant to the review. This search
strategy is an efficient approach for systematic reviews of diagnostic
test accuracy studies (Preston et al. 2015). Unpublished and ongoing
studies were located by searching the International Clinical Trials
Platform.

Published conference abstracts were included if there was enough
information reported to appraise the quality of the study. There were no
language restrictions applied during the search. \textgreater{}The
Cochrane-recommended search strategy combining terms for the `target
condition' and `index test' was used (Mann and Gilbody 2012).

The search strategies used for each data base can be located in
\href{}{file 1}.

The selection of studies was undertaken by two independent reviewers.
Studies were included if reported measures from a ZHF and comparator
device coincided. Studies involving a case control design were excluded
due to potential for overestimation of the intervention performance.
Studies were excluded if conducted on non-human subjects or outside of a
clinical healthcare setting. Studies were excluded if conducted earlier
than the year 2000 as prior studies evaluated outdated technology that
is no longer relevant to current clinical practice.

@@

Information was extracted regarding study characteristics (author, year
of publication, country, design, sample size, clinical setting, numbers
studied and analyses for each outcome), population characteristics
(inclusion and exclusion criteria) and ZHF characteristics (placement of
sensor, timing and methods of measurements). The outcomes that were
extracted included the mean bias (eg, accuracy) and variance (eg, SD,
precision) in temperature measurement between the ZHF sensor and
comparator device. \textgreater{}Information was extracted about how
repeated measurements were handled: (1) analysed each pair of data
separately; (2) treated each pair of data as independent; or (3) used
either analysis of variance or a random effects model as a way to
control for the dependent nature of the repeated measures data (Myles
and Cui 2007).

Using the revised Quality Assessment of Diagnostic Accuracy Studies
(QUADAS-2), each reviewer independently assessed the risk of bias for
the included studies (Whiting et al. 2011). Using guiding questions,
reviewers rated the risk of bias for patient selection, conduct of the
ZHF measurements, conduct of the comparator device measurements, and
timing and flow (eg, timing of ZHF and established core temperature
measurements, dropouts) as ``high'', ``low'' or ``unclear''. We worked
to minimize the risk of publication bias by conducting a comprehensive
search of multiple databases as well as an international clinical trial
registry (Glasziou et al. 2001). \textgreater{}Statistical analyses to
detect reporting bias were not conducted due to lack of validated
methods (Begg 2005). Although some meta-analyses of method comparison
studies have used tests for detecting funnel plot asymmetry (Niven et
al. 2015), simulations have revealed that such tests will result in
publication bias being incorrectly identified too often (Deeks,
Macaskill, and Irwig 2005).

@@

The objective for the meta-analysis was to estimate the population
limits of agreement between temperature measurements from the ZHF and
established comparator devices. \textgreater{}The framework for
meta-analysis of Bland-Altman method comparison studies based on limits
of agreement (LoA) approach was used (Tipton and Shuster 2017).

This method was selected because it parallels the approach used in
primary Bland-Altman studies, whereby an estimate is generated for the
pooled LoAs in the population (not just in the samples studied). The
`population LoA' is more broad than the LoA commonly reported in the
meta-analyses of Bland-Altman studies (Tipton and Shuster 2017). In
these instances, \textgreater{}the pooled LoAs are calculated using
\(\delta\pm2\sqrt{\sigma2+\tau2}\), where \(\delta\) is the average bias
across studies, \(\sigma2\) is the average within-study variation in
differences and \(\tau2\) is the variation in bias across studies.

Estimations of \(\delta\) and \(\sigma2\) were made using a weighted
least-squares model (similar to a random-effects approach), with the
associated SEs estimated using robust variance estimation (RVE). RVE was
used alternatively to model-based SEs as several studies included in the
systematic review used repeated-measures designs without accommodating
for the correlation between measurements (Hedges, Tipton, and Johnson
2010; Tanner-Smith, Tipton, and Polanin 2016; Tipton 2015). We used the
\textgreater{}method-of-moments estimator from ref 16 for the \(\tau2\)
parameter.

In accordance with ref 12, (1) measures of uncertainty were included
\textgreater{}when interpreting the LoA estimates by calculating the
outer \(95\%\) CIs for pooled LoA;

and (2) repeated measurements were adjusted if improperly adjusted for
in \textgreater{}individual studies (by using weights proportional to
the number of participants, not the total number of measurements).

The formulas for the above calculations from ref 12 can be found in
\href{}{file 1}. The R statistical program was used to conduct all
analyses (Team 2017). All data and R code (provided in ref 12) used in
the meta-analyses can be located \href{https://doi.org}{here}.

Prior to conducting the meta-analyses, the results from each study were
converted into a standard format, with bias meaning \(comparator-ZHF\)
measured in \(^\circ C\). In several studies, results were reported for
multiple groups of participants, therefore in the meta-analysis each of
these groups was treated as a separate `comparison'. Other studies
reported multiple sets of results, whereby analyses were conducted
between ZHF and various comparator devices used on the same participant.
These instances were also treated as a separate `comparison' if the
comparator devices were apart of separate meta-analyses groups. One
study reported intraoperative, postoperative and overall results for the
same participants. Only the paired measurements from the overall results
were included in the main and low risk bias analyses, leaving paired
measurements exclusively from the intraoperative and postoperative
timepoints to be included in respective meta-analyses subgroups. The
conventionally cited clinically acceptable agreement between ZHF and
comparator devices is 0.5 \(^\circ C\)
\((ref - most of the studies site this range from Eshragi...)\).
\textgreater{}It was deemed that outer confidence bounds for 95\% LoA
between ECT and ZHF (termed as `population limits of agreement') outside
of these bounds would

be clinically unacceptable. Sensitivity analysis for the primary
meta-analysis was performed based on risk of bias, whereby `unclear risk
of bias' was treated as `high risk' and `high risk of bias' studies were
excluded from analyses. As clinicians would be interested in the
accuracy of ZHF relative to the thermometer devices they use, and within
the clinical setting in which they use it we conducted subgroup analyses
according to the comparator device used (eg, core, peripheral,
nasopharangeal) and clinical setting (eg, intraoperative, intensuve care
unit).

@@

Sixteen studies were eligible for inclusion in this systematic review
(Figure 1).

\includegraphics[width=53.08in]{figure_one}

@@

Using these studies, we conducted 6 different meta-analyses. The first
group compared ZHF to core (eg, arterial, bladder, esophageal, rectal)
temperature measurements and included 22 (73\%) comparisons from 14 (\%)
studies, and enrolled 607 participants with 179,821 paired measurements.
The second group compared ZHF to core temperature measurements in low
risk bias studies studies that scored a low risk of bias in all
categories) and included 10 (\%) comparisons from 5 (\%) studies, and
enrolled 273 participants with 104,294 paired measurements. The third
group compared ZHF to core temperature measurements in ICU patients and
included 7 (23\%) comparisons from 5 (31\%) studies, and enrolled 246
participants with 155,598 paired measurements. The fourth group compared
ZHF to core temperature measurement in intraoperative patients and
included 16 (53\%) comparisons from 10 (63\%) studies, and enrolled 464
participants with 24,223 paired measurements. The fifth group compared
ZHF to nasopharyngeal temperature measurements and included 4 (13\%)
comparisons from 4 (25\%) studies, and enrolled 297 participants with
109,819 paired measurements. The sixth group compared ZHF to sublingual
temperature measurement and included 2 (7\%) comparisons from 2 (13\%)
studies, and enrolled 107 participants with 22731 paired measurements.
The studies included in this systematic review evaluated a ZHF
temperature monitoring system manufactured by 3M. Previously known as
the SpotOn Temperature Monitoring System, the 3M ZHF device is now
referred to as the Bair Hugger Temperature Monitoring System. All
comparisons reported adherence to the ZHF device manufacturer
instructions and placed the sensor on the forehead of the participants.
High risk of bias was associated with patient selection for 13 (43\%)
comparisons from 9 (56\%) studies, conduct of ZHF and comparator
measurements in 10(33\%) comparisons from 8 (50\%) studies and 12 (40\%)
comparisons 9 (30\%) studies, respectively (mostly due to ZHF
measurements being taken with knowledge of the comparator measurements
and vice versa) and participant flow for 13 (43\%) studies. In 19 (50\%)
comparisons from 8 (50\%) studies, the authors had declared a conflict
of interest or receipt of funding or equipment from the manufacturer of
the ZHF device under evaluation.

@@

@@

\leavevmode\hypertarget{ref-niven2015accuracy}{}%
Niven, Daniel J, Jonathan E Gaudet, Kevin B Laupland, Kelly J Mrklas,
Derek J Roberts, and Henry Thomas Stelfox. 2015. ``Accuracy of
Peripheral Thermometers for Estimating Temperature: A Systematic Review
and Meta-Analysis.'' \emph{Annals of Internal Medicine} 163 (10).
American College of Physicians: 768--77.

>>>>

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

@@

    \titlerunning{Accuracy of zero-heat-flux temperature monitoring}

\author{  Aaron Conway \and  }

    \authorrunning{ Conway }

\institute{
        Aaron Conway \at
     Peter Munk Cardiac Centre, UHN \& Lawrence S. Bloomberg Faculty of
 Nursing, University of Toronto \\
     \email{\href{mailto:aaron.conway@utoronto.ca}{\nolinkurl{aaron.conway@utoronto.ca}}}  %  \\
%             \emph{Present address:} of F. Author  %  if needed
    \and
    }

@@

Zero-heat-flux thermometry is an alternative non-invasive method that
allows for continuous monitoring of core temperature. Originally
developed in the 1970s, zero-heat-flux technology was recently
implemented in the 3M SpotOn (3M, St Paul, MN), a single-use, disposable
sensor \((ref)\). In practice, the zero-heat-flux sensor, such as the 3M
SpotOn, is placed on the lateral surface of the forehead and initially
warmed to equilibrate the temperature of the skin surface to the
underlaying core tissues\((ref)\). Equipped with a thermal insulator,
the zero-heat-flux sensor eliminates heat loss to the environment to
allow for changes in core temperature to be directly reflected by a
change in skin surface temperature \((ref)\).

The agreement between zero-heat-flux thermometers and core as well as
other peripheral thermometers has been investigated in several studies
conducted within clinical settings. Appraisal of these studies and
synthesis of the results would aid clinicians in deciding the
appropriate circumstances in which zero-heat-flux thermometers may be
used. We aimed to determine if zero-heat-flux thermometers have
clinically acceptable accuracy and precision relative to established
core and peripheral temperature measurement devices. Accuracy is defined
as the average difference between temperature measurements from the
zero-heat-flux and comparator device and precision as the variance
(standard deviation) in the differences.

@@

A systematic review was conducted in accordance with a predetermined
protocol. The protocol was submitted to PROSPERO for registration but
due to delays in processing we had started data extraction by the time
it was reviewed. Therefore, the protocol did not meet the requirements
for registration. A copy of the submitted protocol can be accessed
\href{insert\%20link}{here}. The primary comparison for this review was
temperature measured from a zero-heat-flux thermometer versus a
temperature measured from a core site, which we defined as arterial,
esophaeal, bladder and rectal sites. Secondary comparisons were made
between zero-heat-flux thermometers and temperatures taken at peripheral
sites.

\hypertarget{inclusion-criteria}{%
\subsection{Inclusion criteria}\label{inclusion-criteria}}

Observational studies that reported temperature measurements from a
zero-heat-flux thermometer and comparator thermometer were included.
Studies involving a case control design were excluded due to potential
for overestimation of the intervention performance. Studies were
excluded if conducted on non-human subjects or outside of a clinical
healthcare setting. No publication date restrictions were
applied.Published conference abstracts were included if there was enough
information reported to appraise the quality of the study. There were no
language restrictions applied during the search.

@@

Published studies were found by searching Medline and EMBASE from
January 2000 to July 2019. The Cochrane-recommended search strategy
combining terms for the `target condition' and `index test' was used
(Mann and Gilbody 2012). This search strategy is an efficient approach
for systematic reviews of diagnostic test accuracy studies.(Preston et
al. 2015) We also conducted forward citation searching, by using Google
Scholar to search the citations of the first article published on the
accuracy of zero-heat-flux thermometers. The search strategies used for
each data base can be located in \href{}{file 1}. Selection of studies
was undertaken independently by two reviewers using
\href{https://www.covidence.org}{Covidence}.

@@

Information was extracted regarding study characteristics (author, year
of publication, country, design, sample size, clinical setting, numbers
studied and analyses for each outcome), population characteristics
(inclusion and exclusion criteria) and temperature measurement
characteristics (placement of sensor, timing and methods of
measurements). The outcomes that were extracted included the mean bias
(eg, accuracy) and variance (eg, SD, precision) in temperature
measurement between the zero-heat-flux and comparator thermometers. We
also extracted information about how repeated measurements were handled.
In particular we assessed whether studies: (1) analysed each pair of
data separately; (2) treated each pair of data as independent; or (3)
used either analysis of variance or a random effects model as a way to
control for the dependent nature of the repeated measures data (Myles
and Cui 2007).

Two reviewers independently assessed the risk of bias for the included
studies using the revised Quality Assessment of Diagnostic Accuracy
Studies (QUADAS-2) (Whiting et al. 2011). Reviewers rated the risk of
bias for patient selection, conduct of the zero-heat-flux measurements,
conduct of the comparator thermometer measurements, and timing and flow
(eg, timing of ZHF and established core temperature measurements,
dropouts) as ``high'', ``low'' or ``unclear'' risk of bias. We worked to
minimize the risk of publication bias by conducting a comprehensive
search of multiple databases as well as an international clinical trial
registry (Glasziou et al. 2001). Statistical approaches for detection of
reporting bias were not conducted due to lack of validated methods (Begg
2005). ASimulations have revealed that tests for detecting funnel plot
asymmetry will result in publication bias being incorrectly identified
too often (Deeks, Macaskill, and Irwig 2005).

@@

The objective for the meta-analysis was to estimate the population
limits of agreement between temperature measurements from the ZHF and
established comparator thermometers. A framework for meta-analysis of
Bland-Altman method comparison studies developed by Tipton and Shuster
(2017) based on limits of agreement approach was used. This method was
selected because it parallels the approach used in primary Bland-Altman
studies, whereby an estimate is generated for the pooled LoAs in the
population (not just in the samples studied). The `population LoA' is
more broad than the LoA commonly reported in the meta-analyses of
Bland-Altman studies (Tipton and Shuster 2017). Pooled limits of
agreement were calculated using \(\delta\pm2\sqrt{\sigma2+\tau2}\),
where \(\delta\) is the average bias across studies, \(\sigma2\) is the
average within-study variation in differences and \(\tau2\) is the
variation in bias across studies.

Estimations of \(\delta\) and \(\sigma2\) were made using a weighted
least-squares model (similar to a random-effects approach), with the
associated standard errors estimated using robust variance estimation.
Robust variance estimation was used alternatively to model-based
standard errors as some studies included in the systematic review used
repeated-measures designs without accommodating for the correlation
between measurements (Hedges, Tipton, and Johnson 2010; Tanner-Smith,
Tipton, and Polanin 2016; Tipton 2015). We used the method-of-moments
estimator from DerSimonian and Laird (1986) for the \(\tau2\) parameter.

Measures of uncertainty were included in our meta-analyses by
calculating the outer \(95\%\) confidence intervals for pooled limits of
agreement. We also accounted for repeated measurements if they were not
properly adjusted for in individual studies. This was achieved by using
weights proportional to the number of participants, not the total number
of measurements. The R statistical program was used to conduct all
analyses (Team 2017). All data and R code used in the meta-analyses can
be located \href{}{here}.

Prior to conducting the meta-analyses, the results from each study were
converted into a standard format, with bias meaning
\(comparator thermometer-ZHF thermometer\) measured in \(^\circ C\). In
several studies, results were reported for multiple groups of
participants, therefore in the meta-analysis each of these groups was
treated as a separate `comparison'. Other studies reported multiple sets
of results, whereby analyses were conducted between ZHF and various
comparator devices used on the same participant. These instances were
also treated as a separate `comparison' if the comparator devices were
apart of separate meta-analyses groups. One study reported
intraoperative, postoperative and overall results for the same
participants. Only the paired measurements from the overall results were
included in the main and low risk bias analyses, leaving paired
measurements exclusively from the intraoperative and postoperative
timepoints to be included in respective meta-analyses subgroups.

The conventionally cited clinically acceptable agreement between ZHF and
comparator devices is 0.5 \(^\circ C\)
\((ref - most of the studies site this range from Eshragi...)\). It was
deemed that outer confidence bounds for 95\% LoA between zero-heat-flux
and core temperature measurements (termed as `population limits of
agreement') outside of these bounds wouldbe clinically unacceptable.

A sensitivity analysis for the primary comparison (ZHF versus
temperature measurement at core site) was performed based on risk of
bias, whereby `unclear risk of bias' was treated as `high risk' and
`high risk of bias' studies were excluded from analyses. As clinicians
would be interested in the accuracy of ZHF relative to the thermometer
devices they use, and within the clinical setting in which they use it,
we conducted subgroup analyses according to the comparator device used
(either core, sublingual or nasopharangeal) and clinical setting (either
intraoperative or intensive care unit).

@@

Fifteen studies were included in this systematic review (Figure 1). Two
studies reported only in abstract form were not included and assigned as
`studies awaiting classification' because there was insufficient
information provided.

\begin{figure}

{\centering \includegraphics[width=4.86in]{zhf_manuscript_files/figure-latex/fig-1-1}

}

\caption{PRISMA Flow Diagram}\label{fig:fig-1}
\end{figure}

@@

The primary comparison of zero-heat-flux versus core temperature
measurements (eg, arterial, bladder, esophageal or rectal) consisted of
19 comparisons from 13 individual studies. In total, data from 576
participants with 179821 paired measurements were included in this
comparison (two studies did not report the total number of measurements
included in their analysis). The sensitivity analysis for the primary
comparison with only studies that were judged as low risk of bias across
all domains included 10 comparisons from 5 studies that enrolled 273
participants with 104294 paired measurements (two studies did not report
the total number of measurements included in their analysis). The third
group compared zero-heat-flux to core temperature measurements in ICU
patients and included 7 (23\%) comparisons from 5 (31\%) studies, and
enrolled 246 participants with 155,598 paired measurements. The fourth
group compared ZHF to core temperature measurement in intraoperative
patients and included 16 (53\%) comparisons from 10 (63\%) studies, and
enrolled 464 participants with 24,223 paired measurements. The fifth
group compared ZHF to nasopharyngeal temperature measurements and
included 4 (13\%) comparisons from 4 (25\%) studies, and enrolled 297
participants with 109,819 paired measurements. The sixth group compared
ZHF to sublingual temperature measurement and included 2 (7\%)
comparisons from 2 (13\%) studies, and enrolled 107 participants with
22731 paired measurements. The studies included in this systematic
review evaluated a ZHF temperature monitoring system manufactured by 3M.
Previously known as the SpotOn Temperature Monitoring System, the 3M ZHF
device is now referred to as the Bair Hugger Temperature Monitoring
System. All comparisons reported adherence to the ZHF device
manufacturer instructions and placed the sensor on the forehead of the
participants. High risk of bias was associated with patient selection
for 13 (43\%) comparisons from 9 (56\%) studies, conduct of ZHF and
comparator measurements in 10(33\%) comparisons from 8 (50\%) studies
and 12 (40\%) comparisons 9 (30\%) studies, respectively (mostly due to
ZHF measurements being taken with knowledge of the comparator
measurements and vice versa) and participant flow for 13 (43\%) studies.
In 19 (50\%) comparisons from 8 (50\%) studies, the authors had declared
a conflict of interest or receipt of funding or equipment from the
manufacturer of the ZHF device under evaluation.

@@

\leavevmode\hypertarget{ref-dersimonian1986meta}{}%
DerSimonian, Rebecca, and Nan Laird. 1986. ``Meta-Analysis in Clinical
Trials.'' \emph{Controlled Clinical Trials} 7 (3). Elsevier: 177--88.

@@
